---
title: "Getting Started with BDR Utils"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with BDR Utils}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(BDRUtils)
```

## Exclusivity Table Calculation

The `exclusivity_table` function in the `BDRUtils` package is a powerful tool to calculate exclusivity statistics for specified variables in a given data frame. It is particularly useful when you have a dataset with dummy variables (encoded as "Yes" and NA) and want to analyze the exclusivity of these variables with respect to an 'Area_sq_Km' column, representing the shared area in square kilometers.

### Usage

To use the `exclusivity_table` function, you need to provide two arguments:
 
- `Vars`: A character vector of variable names for which exclusivity should be calculated.
- `DF`: A data frame containing the dummy variables and the 'Area_sq_Km' column.

### Output

The function returns a data frame summarizing exclusivity statistics for the specified variables. The resulting data frame includes the following columns:

- `Variable`: The variable name for which exclusivity statistics are calculated.
- `Total`: The total shared area in square kilometers for each variable.
- `Exclusive`: The area that exclusively belongs to the specified variable (non-overlapping with other variables).
- `Non_exclusive`: The area that overlaps with other variables.

### Example

Let's consider a sample data frame 'LT':
 
```{r}
 LT <- data.frame(
   planned_strictly_protected = c("Yes", "Yes", "Yes", "Yes", NA, NA, NA, NA),
   strictly_protected = c("Yes", "Yes", NA, NA, "Yes", "Yes", NA, NA),
   sea_natural_areas = c("Yes", NA, "Yes", NA, "Yes", NA, "Yes", NA),   Area_sq_Km = c(4163.9369, 0.0043, 136.5007, 0.0001, 2153.2234, 23.0762, 24540.5325, 327940.2363)
 )
```

as seen in the following table

```{r, echo = F}
knitr::kable(LT)
```

You can calculate the exclusivity table for the 'strictly_protected' and 'sea_natural_areas' variables using the following code:

```{r}
result <- exclusivity_table(DF = LT, Vars = c("strictly_protected", "sea_natural_areas"))
```

The resulting data frame will provide insights into the exclusive and non-exclusive areas for each specified variable as seen in the following table.

```{r, echo = F}
knitr::kable(result, caption = "Resulting table for exclusivity_table function", digits = 2)
```

