---
title: "Getting Started with BDR Utils"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with BDR Utils}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(BDRUtils)
```

## Exclusivity Table Calculation

The `exclusivity_table` function in the `BDRUtils` package is a powerful tool to calculate exclusivity statistics for specified variables in a given data frame. It is particularly useful when you have a dataset with dummy variables (encoded as "Yes" and NA) and want to analyze the exclusivity of these variables with respect to an 'Area_sq_Km' column, representing the shared area in square kilometers.

### Usage

To use the `exclusivity_table` function, you need to provide two arguments:
 
- `Vars`: A character vector of variable names for which exclusivity should be calculated.
- `DF`: A data frame containing the dummy variables and the 'Area_sq_Km' column.

### Output

The function returns a data frame summarizing exclusivity statistics for the specified variables. The resulting data frame includes the following columns:

- `Variable`: The variable name for which exclusivity statistics are calculated.
- `Total`: The total shared area in square kilometers for each variable.
- `Exclusive`: The area that exclusively belongs to the specified variable (non-overlapping with other variables).
- `Non_exclusive`: The area that overlaps with other variables.

### Example

Let's consider a sample data frame 'LT':
 
```{r}
LT <- data.frame(
  planned_strictly_protected = c("Yes", "Yes", "Yes", "Yes", NA, NA, NA, NA),
  strictly_protected = c("Yes", "Yes", NA, NA, "Yes", "Yes", NA, NA),
  sea_natural_areas = c("Yes", NA, "Yes", NA, "Yes", NA, "Yes", NA), Area_sq_Km = c(4163.9369, 0.0043, 136.5007, 0.0001, 2153.2234, 23.0762, 24540.5325, 327940.2363)
)
```

as seen in the following table

```{r, echo = F}
knitr::kable(LT)
```

You can calculate the exclusivity table for the 'strictly_protected' and 'sea_natural_areas' variables using the following code:

```{r}
result <- exclusivity_table(DF = LT, Vars = c("strictly_protected", "sea_natural_areas"))
```

The resulting data frame will provide insights into the exclusive and non-exclusive areas for each specified variable as seen in the following table.

```{r, echo = F}
knitr::kable(result, caption = "Resulting table for exclusivity_table function", digits = 2)
```

## Write Cloud Optimized Geotiff (COG)

The `write_cog` function in the `BDRUtils` package allows you to save a SpatRaster object as a Cloud Optimized Geotiff (COG). COGs are geospatial files optimized for efficient storage and retrieval, particularly useful for cloud-based geospatial data platforms.

### Usage

To use the `write_cog` function, you need to provide two arguments:

- `SpatRaster`: A SpatRaster object (class SpatRaster) representing the raster data you want to save as a COG.
- `Name`: The desired name for the COG file, including the ".tif" extension.

### Output

The function returns a Cloud Optimized Geotiff file saved at the specified location. This file is optimized for efficient cloud storage and retrieval.

### Example

Let's consider a simple example:

```{r}
# Load required libraries if not already loaded
# library(terra)

# Create a sample SpatRaster
r <- terra::rast(nrows = 5, ncols = 5, vals = 1:25)

# Save the SpatRaster as a COG
write_cog(SpatRaster = r, Name = "test.tif")
```

In this example, we create a SpatRaster r, specify a filename for the COG file ("test.tif"), and then use the write_cog function to save the SpatRaster as a COG.

### Why Use COGs?

Cloud Optimized Geotiffs (COGs) are designed for optimized cloud-based geospatial data storage. They allow for efficient reading and streaming of geospatial data from cloud storage platforms such as AWS S3, Google Cloud Storage, and more. COGs are widely adopted in the geospatial community for sharing and analyzing large geospatial datasets efficiently.

## Subclass classification

There are several functions available to classify land use into specific categories. Each function uses specific criteria to determine the classification. Here's a breakdown of each function:

### OpenGenerator

- **Logic**: This function checks a DataFrame and classifies rows as "Open" habitat based on specific criteria.
  
- **Definition**: If the value in the `p3_klit` column (defining habitat types within paragraph 3) is one of the specified habitat types ("Eng", "Hede", "Mose", "Overdrev", "Strandeng", "Klit"), and it is not NA, and the row is not classified as "Untouched" forest (`Urort_Skov` is NA), and it is not classified as "Forest" (`Forest` is NA), and it is not classified as "Intensive Agriculture" (`markblokkort` is NA), then it is classified as "Open" in the `Subclass` column of the DataFrame.

### SoGenerator

- **Logic**: This function checks a DataFrame to classify rows as "So" (lake) habitat based on specific criteria.
  
- **Definition**: If the value in the `p3_klit` (paragraph 3) column is "Sø" and it is not NA, then it is classified as "So" (lake) in the `Subclass` column of the DataFrame.

### SkovnaturGenerator

- **Logic**: This function checks a DataFrame to classify rows as "Skovnatur" (natural forest) habitat based on specific criteria.
  
- **Definition**: If the row meets one of the following conditions:
  - It is classified as "Untouched" forest (`Urort_Skov` is not NA) OR
  - It is classified as "Forest" (`Forest` is not NA), but not as "Untouched" forest (`Urort_Skov` is NA), and it is either in a "National Park" (`National_Parks` is not NA), "Private foundation" (`Fondsejede` is not NA), or falls under the specified habitat types in the `p3_klit` column, then it is classified as "Skovnatur" in the `Subclass` column of the DataFrame.

### DrevetSkovGenerator

- **Logic**: This function checks a DataFrame to classify rows as "Drevet Skov" (managed forest) habitat based on specific criteria.

- **Definition**: If the row meets the following conditions:
  - It is classified as "Forest" (`Forest` is not NA) AND
  - It is not classified as "Untouched" forest (`Urort_Skov` is NA) AND
  - It is not classified as part of a "Private foundation" (`Fondsejede` is NA) AND
  - It is not classified as "National Park" (`National_Parks` is NA) AND
  - It is not classified as "Intensive Agriculture" (`markblokkort` is NA), then it is classified as "Drevet Skov" in the `Subclass` column of the DataFrame.

### PGR_in_P3Generator

- **Logic**: This function checks a DataFrame to classify rows as "PGR_in_P3" (Permanent grasslands in paragraph 3) habitat based on specific criteria.

- **Definition**: If the row meets the following conditions:
  - The value in the `p3_klit` column is not NA AND
  - The value in the `markblokkort` column is "PGR" AND
  - The value in the `p3_klit` column is not "Soer" (not a lake), then it is classified as "PGR_in_P3" in the `Subclass` column of the DataFrame.

### PGR_out_of_P3Generator

- **Logic**: This function checks a DataFrame to classify rows as "PGR_out_of_P3" (Permanent grasslands outside of paragraph 3) habitat based on specific criteria.

- **Definition**: If the row meets the following condition:
  - The value in the `p3_klit` column is NA (not in paragraph 3) AND
  - The value in the `markblokkort` column is "PGR" (permanent grassland), then it is classified as "PGR_out_of_P3" in the `Subclass` column of the DataFrame.

### INT_AGG_Generator

- **Logic**: This function checks a DataFrame to classify rows as "INT_AGG" (intensive agriculture) habitat based on specific criteria.

- **Definition**: If the row meets the following conditions:
  - The value in the `markblokkort` column is "INT_AGG" (intensive agriculture) AND
  - The value in the `p3_klit` column is not "Soer" (not a lake), then it is classified as "INT_AGG" in the `Subclass` column of the DataFrame.

These functions provide a systematic way to classify rows in your DataFrame based on various habitat and land-use criteria. They help you generate specific habitat classes in the `Subclass` column, allowing for more structured analysis of your data.

It's important to note that these categories might not be exclusive of each other, so the order in which you apply these functions can affect the total area for each subclass.
